\documentclass{article}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{indentfirst}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,calc}
\pgfplotsset{compat=1.18}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{problem}{Problem}
\newtheorem{solution}{Solution}
\newtheorem*{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{method}{Method}
\newtheorem{principle}{Principle}
\newcommand{\Span}{\operatorname{span}}
\newenvironment{amatrix}[1]{%
  \left(\begin{array}{@{}*{#1}{c}|c@{}}
}{%
  \end{array}\right)
}
\reversemarginpar
\hypersetup{colorlinks=true,}
\begin{document}

%------- Title page -----------
\title{MATH 323: Probability}
\author{W.R.YCH.}
\date{May \(4^{th}\), 2026}
\setcounter{Maxaffil}{0}
\renewcommand\Affilfont{\itshape\small}
\maketitle

%------- Abstract -----------
\noindent\rule{\textwidth}{0.4pt}
\thispagestyle{empty}
\begin{abstract}
\end{abstract}
\noindent\rule{\textwidth}{0.4pt}
\clearpage

%------- Table of Contents -----------
\thispagestyle{empty}
{
  \hypersetup{linkcolor=black}
  \tableofcontents
}
\clearpage

%------- Introduction -----------
\setcounter{page}{1}
\section{Introduction}
These notes were written for MATH 323 (Probability) at McGill University, Summer 2026. The course is taught by Magid Sabbagh and meets in Steward Biology Building S1/3 (S1/3) every Monday, Tuesday, Wednesday, and Thursday from 8:35\,am to 10:55\,am.
\subsubsection*{Grading scheme:}
\vspace{-2em}
\begin{align}
    20\%\ \text{assignments} + 5\% \text{Multiple Choice Quiz} + 25\%\ \text{midterm} + 50\%\ \text{final}\\
    \text{or}\\
    20\%\ \text{assignments} + 5\% \text{Multiple Choice Quiz} + 75\%\ \text{final}
\end{align}
whichever is higher. The book used for this class is: "Mathematical statistics with application, 7th edition." Midterm will be an 1h30 long and will have all 2h30 hours to complete it in class.

% ============================================================
\section{Prerequisite Knowledge}
% ============================================================
\marginpar{May 4, 2026}
Basic Set Theory:
\begin{itemize}
    \item A set is a well defined collection of objects, defined in a way such that we can determine wether or not an element $x$ belongs to the set.
    \item If $A$ is a set, we say that $a\in A$, if a is an element of A.
    \item If A is a set, we say that $a \notin A$ if a is not an element of A.
    \item We specify a set either by listing its elements, if possible on by a property that defines the set.
    \item In general, we will be working in a universal set $\Omega$.
    \item Universal set is a set in which all elements belong, like the complex set.
    \item Operations on sets, let $\Omega$ be a universal set, $A, B \subseteq \Omega$. Then we can define the following operations on sets:
    \begin{itemize}
        \item Subset: $A \subseteq B$ if $x \in A$ implies $x \in B$.
        \item Equality: $A = B$ if $A \subseteq B$ and $B \subseteq A$.
        \item Proper subset: $A \subsetneq B$ if $A \subseteq B$ and $A \neq B$.
        \item Union: $A \cup B = \{x : x \in A \text{ or } x \in B\}$
        \item Intersection: $A \cap B = \{x : x \in A \text{ and } x \in B\}$
        \item Difference: $A \setminus B = \{x : x \in A \text{ and } x \notin B\}$
        \item Complement: $A^c = \{x : x \in \Omega \text{ and } x \notin A\}$
        \item Singleton: $\{x\} = \{y : y = x\}$, the set that contains only the element $x$.
    \end{itemize}
\end{itemize}
The complement of $A \subseteq \Omega$ is denoted by $A^c$ or $\overline{A}$ and is defined to be $A^c = \{x\in \Omega: x\notin A\}$.
The empty set is denoted by $\emptyset$.
\begin{proposition}
    $A \cup A^c = \Omega$, $A\cap A^c = \emptyset$, $A\cup \emptyset = A$, $A\cap \emptyset = \emptyset$.
\end{proposition}
\begin{proposition}
    Let A,B,C be sets:
    \begin{enumerate}
        \item Commutativity: $A\cup B = B\cup A$.
        \item Associativity: $A\cup(B\cup C)= (A\cup B)\cup C$, $A\cap (B\cap C) = (A\cap B)\cap C$.
        \item Distributivity: $A\cup (B\cap C) = (A\cup B)\cap (A\cup C)$, $A\cap (B\cup C) = (A\cap B)\cup (A\cap C)$.
    \end{enumerate}
\end{proposition}
\begin{proof}
    $A\cap(B\cup C) = (A\cap B)\cup(A\cap C)$, $A\cap(B\cup C) \subset (A\cap B)\cup (A\cap C)$.
\end{proof}
Let $x\in A\cap(B\cup C)$. Then $x\in A$ and $x\in B\cup C$. Then $x\in A$ and $x$ belongs to atleast B and C. If $x\in A$ and $x\in B$ then $x\in A\cap(B\cup C)$?
If $x\in A$ and $x\in C$, then $A\cap C \subset (A\cap B)\cup(A\cap C)$. Hence $A\cap (B \cup C)\subset(A\cap C)$. We now show that $A\cap B)\cup(A\cup C) \subset A\cap(B\cup C)$. Let $x\in (A\cap B)\cup(A\cap C)$. If $x \in A\cap B$ then $x\in A$ and $x\in B \subseteq B\cup C$. If $x\in A\cap C$, then $x\in A$ and $x\in C\subseteq B\cup C$. So $x \in A\cap (B\cup C)$. Hence $(A\cap B)\cup(A\cap C) \subset A\cap(B\cup C)$.
De morgans law:
\begin{enumerate}
    \item $(A\cup B)^c = A^c \cap B^c$.
    \item $(A\cap B)^c = A^c \cup B^c$.
\end{enumerate}
Let $A_1, A_2, A_3, \ldots$ be subsets of $\Omega$. Let $n\in \mathbb{N}: \cap_{i=1}^n A_i = A_1 \cap A_2 \cap \ldots \cap A_n = \{x\in \Omega: x\in A_1, \ldots, x\in A_n\}$.
$\cap_{i=1}^\infty A_i = \{x\in \Omega: x\in A_i \forall i \in \mathbb{N}\}$.
$\cup_{i=1}^\infty A_i = \{x\in \Omega: x\in A_{i_0} \text{ for some } i_0 \in \{1,\ldots,n\}\} = \{x\in \Omega: x\in \text{at least one of } A_1,\ldots,A_n\}$.
\begin{proposition}
    $(\cup_{i=1}^\infty A_i)^c = \cap_{i=1}^\infty A_i^c$ and $(\cap_{i=1}^\infty A_i)^c = \cup_{i=1}^\infty A_i^c$.
\end{proposition}
\begin{problem}
    In a class of 101 students. 20 students play hockey. 30 students like math. 5 students like math and play hockey. How many students, (1) do not play hockey, (2) do not like math, (3) do not like math and do not play hockey, (4) play hockey but do not like math. Answer (1): $101 - 20 = 81$ students do not play hockey. (2) $101 - 30 = 71$ students do not like math. (3) $101 - (20 + 30 - 5) = 56$ students do not like math and do not play hockey. (4) $20 - 5 = 15$ students play hockey but do not like math.
\end{problem}
We define $A\setminus B = \{x\in \Omega: x\in A and x\notin B\} = A\cap B^c = A\setminus (A\cap B)$.
\begin{remark}
    Write $|A|$ (cardinal of A) for the number of elements in A. Then $|A\cup B| = |A| + |B| - |A\cap B|$.
\end{remark}
\begin{proof}
    $|A\cup B| = x+y+z, |A| = x+y, |B| = y+z, |A\cap B| = y$. Then $|A\cup B| = |A| + |B| - |A\cap B|$.
\end{proof}

\marginpar{May 5, 2026}
\begin{example}
    1) Let $\Omega$ be a set. $\mathbb{F} = \{\phi, \Omega\}$ is a $\sigma$-algebra on $\Omega$.\\
    2) Let $\Omega$ be a set. Then the set of all subset of $\Omega$ is a $\sigma$-algebra on $\Omega$. (The power set of $\Omega$ denoted by $Z^\Omega$).\\
    3) Let $\omega$ be a set. Define $\mathbb{F} = \{A \subseteq \Omega. \text{A is countable on } A^c \text{is countable}\}$. Then $\mathbb{F}$ is a $\sigma$-algebra on $\Omega$.
    4) Suppose $A\subseteq \Omega$. $\mathbb{F} = \{\phi, A, A^c, \Omega\}$ is a $\sigma$-algebra.
\end{example}
\begin{example}
    Flip a coin. $\Omega = \{H, T\}$. $2^n = \{\phi, \{H\},\{T\},\{H,T\}\}$.
\end{example}
\begin{definition}
    Let $\Omega$ be the sample space (...) to an experiment and $\mathbb{F}$ a $\sigma$-algebra of subsets of $\Omega$. (...)
    \begin{enumerate}
        \item $P(A) \geq 0$ for all $A \in \mathbb{F}$.
        \item $P(\Omega) = 1$.
        \item If $A_1, A_2, \dots$ is a collection of pairwise disjoint subsets of $\Omega$ that belong to F then are sets in $\mathbb{F}$ such that $A_i \subseteq A_{i+1}$ for all $i$, then $P(\cup_{i=1}^\infty A_i) = \sum_{i=1}^\infty P(A_i)$. (Pairwise disjoint $A_k \cap A_l = \phi$ if $k \neq l$)
    \end{enumerate}
    Properties:
    \begin{enumerate}
        \item If $A \subseteq B$ then $P(A) \leq P(B)$. Proof: $B = A \cup (B\setminus A)$ and $P(B) = P(A) + P(B\setminus A)$ since $A$ and $B\setminus A$ are disjoint. $P(B) = P(A) + P(B\setminus A) \geq P(A)$ since $P(B\setminus A) \geq 0$.
    \end{enumerate}
\end{definition}
\begin{corollary}
    Let $A \in \mathbb{F}$. (A is an event). Then $P(A) \leq 1$. Proof: $A \subseteq \Omega$, (...) by the previous proof, $P(A) \leq P(\Omega) = 1$.
\end{corollary}
\begin{proposition}
    $P(A^c) = 1 - P(A)$. Proof: $\Omega = A \cup A^c$ and $A \cap A^c = \phi$. $P(\Omega) = P(A \cup A^c) = P(A) + P(A^c) = 1$, so $P(A^c) = 1 - P(A)$.
\end{proposition}
Let $A, B$ be events $(A, B \in \mathbb{F})$ . Then $A \cap B \in \mathbb{F}$ and $P(A\cup B) = P(A) + P(B) - P(A\cap B)$. $A \cup B = A \cup (B \cap A^c)$, $A \cap (B\cap A^c) = \phi$, $P(A\cup B) = P(A) + P(B\cap A^c)$.
\begin{lemma}
    $P(B) = P(A^c\cap B) + P(A\cap B)$ and $P(A \cap B) = P(B) - P(A^c\cap B)$, $B = (A\cap B)\cup (A^c\cap B)$, $A\cap B$, $A^c \cap B$ are disjoint $P(B) = P(A\cap B)+P(A^c \cap B)$ and hence $P(A\cap B) = P(B) - P(A\cap B)$ and hence $P(A\cup B) = P(A) + P(B) - P(A\cap B)$.
\end{lemma}
The axioms define a probability, require THE countable additivity for piecewise disjoint sets IF $\{A\}^{\infty}_{i=1}$ are a collection of events such that $A_k \cap A_l = \phi$ if $k\neq l$ then $P(\cup_{i=1}^\infty A_i) = \sum_{i=1}^\infty P(A_i)$. If $\{A\}^{\infty}_{i=1}$ are not disjoint, how can we relate $P(\cup_{i=1}^\infty A_i)$ and $\sum_{i=1}^\infty P(A_i)$? 
\begin{proposition}
    Let $\{A\}_{i=1}^\infty$ be a collection of events. Then $P(\cup_{i=1}^\infty A_i) \leq \sum_{i=1}^\infty P(A_i)$. Proof: define $A^\star_i = A_i$ and $A_2^\star = A_2 \setminus A_1$, $A_3^\star = A_3 \setminus (A_1 \cup A_2)$, $\dots$, $A_n^\star = A_n \setminus (\cup_{i=1}^{n-1} A_i)$, $\dots$, for $n\geq 2$.
\end{proposition}
First, $\{A\}_{i=1}^\infty$ are pairwise disjoint. To show this, let $k\neq l$ be positive integers. We need to show that $A_k^\star \cap A_l^\star = \phi$. Assume that $l > k$ and suppose$A_k^\star \cap A_l^\star \neq \phi$. Let $x \in A_k^\star \cap A_l^\star$. Then $x \in A_l^\star$ and $x \in A_k^\star$. Since $x \in A_l^\star$ and $x \in A_k^\star$. Since $x \in A_l^\star$ then $x \in A_l$ but $x\notin \cup_{i=1}^{l-1} A_i$. Since $k<l$ then $x\notin A_k$. Futhermore, since $x\in A_k^\star$ , then $x\in A_k$, we have here a contradiction. Hence $\{A_i^\star\}_{i=1}^\infty$ are piecewise disjoint. Then next step is to show that $\cup_{i=1}^\infty A_i = \cup_{i=1}^\infty A_i^\star$. First, note that since $A_i^\star \subset A_i$ for all $i\in\mathbb{N}$. Then $\cup_{i=1}^\infty A_i^\star \subset \cup_{i=1}^\infty A_i$. Now, we show that $\cup_{i=1}^\infty A_i \subset \cup_{i=1}^\infty A_i^\star$. Suppose $x\in \cup_{i=1}^\infty A_i$. Then, let $i_0$ be the smallest index integer such that $x \in A_{i_0}$. In particular, $x\notin A_1, x\notin A_2,\ldots, x\notin A_{i_0-1}$. Then $x\in A_{i_0}^\star \subset \cup_{i=1}^\infty A_i^\star$. Therefore, $\cup_{i=1}^\infty A_i^\star \subset \cup_{i=1}^\infty A_i^\star$. Then $\cup_{i=1}^\infty A_i^\star = \cup_{i=1}^\infty A_i$. $P(\cup_{i=1}^\infty A_i) = P(\cup_{i=1}^\infty A_i^\star) = \sum_{i=1}^\infty P(A_i^\star) \leq \sum_{i=1}^\infty P(A_i)$, since $A_i^\star \subset A_i$.
\begin{corollary}
    Let $A_1, A_2, \ldots, A_n$ be events. Then $P(\cap_{i=1}^n A_i) \geq \sum_{i=1}^n P(A_i) - (n-1)$. Proof: $P(\cap_{i=1}^n A_i) = 1 - P((\cap_{i=1}^n A_i)^c) = 1- P(\cup_{i=1}^n A_i^c) \geq 1 - \sum_{i=1}^n P(A_i^c) = 1 - (n - \sum_{i=1}^{n}P(A_i)) = \sum_{i=1}^{n}P(A_i) - (n-1)$, since $P(\cup_{i=1}^n A_i^c) \leq \sum_{i=1}^n P(A_i^c)$.
\end{corollary}
\begin{problem}
    Suppose $\{A_i\}_{i=1}^\infty$ are events such that $P(A_i)=1$ for all $i\in \mathbb{N}$ such that $P(\cap_{i=1}^\infty A_i) = 1$.\footnote{He'll prolly ask this on an exam.}
\end{problem}
Independent Events: Events A and B are said to be independent if $P(A \cap B) = P(A)P(B)$.
(...)
Finite sample spaces and equally likely outcomes
\begin{example}
    We roll a dce twice (tolls are assumed to be independent) and we write down the number obtained at each roll. The sample space $\Omega = \{(a,b), a,b \in \{1,2,3,4,5,6\}\}$. $|\Omega| = 36$. Since the two rolls are independent, the probability of $\{(a,b)\}$ is $1/6 x 1/6 = \frac{1}{36}$. The outcomes are equally likely. Let $\Omega = \{w_1,\ldots,w_n\}$ be a finite sample space with $|\Omega| = N$ we say that the outcomes $w_1,\ldots,w_N$ are equally likely if $P(\{w_1\}) = P(\{w_2\}) = \ldots = P(\{w_N\}) = \frac{1}{N}$. Let $A \subset \Omega$. Since $\Omega$ is finite, then A is finite and $A=\{w_1,w_2,\ldots,2_d\}$ and $|A| = d$, $w_1,\ldots,w_n$ are the distinct $d$ elements of A. $P(A) = \sum_{i=1}^d P(\{w_i\}) =  \frac{1}{N} + \frac{1}{N} + \ldots + \frac{1}{N} = \frac{d}{N} = \frac{|A|}{|\Omega|}$.
\end{example}
\begin{example}
    A fair coin is flipped twice. Suppose the flips are independent. The sample space $\Omega = \{HH, HT, TH, TT\}$ outcomes in $\Omega$ are equally likely with probability $1/4$.
    \begin{enumerate}
        \item Probability of getting at least one head, call this event A. Find P(A): $A = \{HH,HT,TH\}$, $|A| = 3$, $P(A) = \frac{3}{4}$.
    \end{enumerate}
\end{example}
Counting
\begin{theorem}[Fundamental Theorem of Counting]
    With m elements $a_1, \ldots, a_m$ and n elements $b_1, \ldots, b_n$, one can form $m \times n$ distinct pairs $(a,b)$ where $a \in \{a_1,\ldots, a_m\}$ and $b \in \{b_1,\ldots, b_n\}$.
\end{theorem}
\begin{example}
    how many 7 digit telephone numbers can one form? 10 chosen for first, 10 chosen for second, ...,  10 chosen for 7th
    How many 7-digit telephone numbers can one form if the third digit is even: 10 chosen for first, 10 chosen for second, 5 chosen for third, 10 chosen for fourth, ..., 10 chosen for seventh.
\end{example}
Permutations
\begin{example}
    In a class of 100 students, 2 students are chosen. The first one wins 100\$ and the second one wins 75\$. How many choices are there if (1) a student can win twice (sampling with replacement) (2) a student can win at most once (sampling without replacement). (1) 100 choices for first, 100 choices for second, so $100^2$ choices. (2) 100 choices for first, 99 choices for second, so $100 \times 99$ choices.
\end{example}
Permutations: An ordered arrangement of r distinct objects is called a permutation. The number of ways of randomly choosing r ordered objects form $n$ objects without replacement is denoted by $P^n_r$ and is equal to $\frac{n!}{(n-r)!}$. In total, the number of ordering is $n\times (n-1) \times \ldots \times (n-r+1)$. $P^n_r = n\times (n-1) \times \ldots \times (n-r+1) = \frac{n!}{(n-r)!}$.
\begin{example}
    Two students are chosen to form a committee. In here many ways can the choice be done? 100 choices for first one, 99 choices for second one, 2! ways of ordering the two students, so $\frac{100 \times 99}{2!}$ ways.
\end{example}
% ============================================================
\section{Appendix}
% ============================================================


% ============================================================
\section{Solutions}
% ============================================================


% ============================================================
\section{Useful Links}
% ============================================================

\end{document}